### 2017-07-18 manuscript.Rmd
WIP

---
title: "An Introduction to Rocker: Docker Containers for R"
author:
  - name: Carl Boettiger
    affiliation: UC Berkeley 
    address:
    - ESPM Department, University of California,
    - 130 Mulford Hall Berkeley, CA 94720-3114, USA
    email:  cboettig@berkeley.edu
  - name: Dirk Eddelbuettel
    affiliation: Debian and R Projects; Ketchum Trading
    address: 
    - Chicago, IL, USA
    email:  edd@debian.org
abstract: >
  We describe the Rocker project, which provides a widely-used suite of Docker images
  with customized R environments for particular tasks. We discuss how this suite is
  organized and how these tools can increase portability, scaling, reproducibility,
  and convenience of R users and developers.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

<!-- Notes to Dirk: 
     Some of this material may be more appropriate for documentation elsewhere, appearing here in only abbreviated form.  
     To get started, I've just tried to err on the side of comprehensive, particularly since our documentation in places is sparse
     or dated.  Feel free to flag any areas that should be moved in part or whole to documentation.  -->

## Introduction

The Rocker project was launched in October 2014 as a collaboration
between CB & DE to provide high-quality Docker images containing the
R environment \citep{edd2014}. Since that time, the project has seen
both considerable uptake in the community and substantial development
and evolution. Here we seek to document the project's objectives and uses.

### What is Docker?

Docker is a popular open source tool to create, distribute, deploy,
and run software applications using *containers*.  Containers provide a
virtual environment requiring all operating-system components an application needs to run:
code, runtime, system tools, system runtime.  Docker containers are
lightweight as they share the operating system kernel, starting instantly,
using a layered filesystem which minimizes disk footprint and download time,
and are built on open standards that run on all major platforms (Linux, Mac,
Windows), and provide an added layer of security by running an application
in an isolated environment \citep{what-docker}.

## Rocker organization and workflow

The Rocker Project consists of a suite of images built automatically by
and hosted on the Docker Hub, <https://hub.docker.com/r/rocker>.  Source
Dockerfiles, supporting scripts and documentation are hosted on GitHub
under the organization `rocker-org`, <https://github.com/rocker-org>.
The issues tracker and pull requests are used for community input,
disucssions and contributions to these images. The rocker project wiki
provides a place to synthesize community-contributed documentation,
use-cases, and other knowledge about using the Rocker images.

### Images in the Rocker Project

The Rocker Project aims to provide a small core of Docker images that
serve as convenient 'base' images on which other users can build custom
R environments by writing their own Dockerfiles, while also providing a
'batteries included' approach of images that can be used out of the box.
The challenges of balancing diverse needs driven by very different use
cases (discussed in more detail below) against the overarching goals
of creating images that are still sufficiently light-weight, easy to use
and easy to maintain is a difficult art.  The implementation in both
individual Rocker images and image stacks can never perfect that balance
for everyone, but today reflects the considerable community input and
testing over past few years.

All Rocker images are based on the Debian Linux Distribution.
The Debian platform provides a small base image, the well-known `apt`
package management system and rich ecosystem of software libraries,
making it the base image of choice for Docker images, including many
of the "official" images maintained by Docker's own development team.
The Debian platform is also perhaps the best supported Linux platform
within the R community, including an active `r-sig-debian` listserve.
Unfortunately, the relatively long period between stable Debian releases
(roughly two years recently) means that software in the `debian:stable`
releases can lag significantly behind current releases of popular
software, including R.  More recent versions of packages can be found in
the pre-release distribution, `debian:testing`, while the very latest
binary builds can be found on `debian:unstable`.  The Rocker Project
can be largely divided into two stacks which address different needs,
defined by which Debian distribution they are based on.  The first stack
is based on `debian:testing`, the second, more recently introduced stack
is based only on `debian` stable releases.  These stacks have different
aims and thus provide different images, as shown in Tables 1 & 2.

## The `debian:testing`-based images

The `debian:testing` stack aims to makes the most efficient use of
upstream builds -- the precompiled `.deb` binaries provided by the
Debian repositories.  It is both quicker and easier to install software
from binaries, since the package manager (`apt`) manages the necessary
(binary) dependencies and bypasses the time-consuming process of compiling
from source.  Basing this stack on `debian:testing` means that much more
recent versions of commonly-used libraries and compilers are available
as binaries than would be found in a Debian stable release.  In order
to provide access to the most recent avialble binaries, this stack uses
apt-pinning [@apt-pinning] to allow the `apt` package manager to also
install binaries from `debian:unstable`, which represents the most recent,
bleeding edge of packages built for Debian when necessary.  For instance,
the `r-base` image provided by Rocker installs the most recent version of
R as a binary from Debian `unstable`.  Similarly, recent versions of many
popular R packages can also be installed through the package manager,
e.g. `apt-get install r-cran-xml`.  This can be particularly helpful
for packages with external system dependencies (such as `libxml2-dev`
in this example) which cannot be installed from the R console as they 
are system dependencies rather than R packages installed from within R.

As the names `testing` and `unstable` imply, this approach is not
without challenges.  The particular version of any given package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occassionally break an previously working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a the package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`, see Box 1 on apt-pinning.)

Relative to `stable`, the `testing` stack thus offers the following advantages:

1. These Dockerfiles are easy to develop and extend because almost all software can be installed through the package manager
2. These Dockerfiles always install the most recent available software
3. These Dockerfiles can almost always build relatively quickly

and these down-sides:

1. These Dockerfiles require occassional minor maintenance to ensure successful builds. (e.g. changing an installation directive from `unstable` to `testing` or vice versa)
2. The resulting images are inherently dynamic: rebuilding the same Dockerfile months or years apart will generate images with significantly different versions of software installed.
3. The use of apt pinning may be unfamiliar to some users, where the user must ultimately be responsible to ensure compatibility [@apt-pinning].
4. `testing` and `unstable` do not recieve the attention to security updates from the Debian security team.
<!-- DE: Are you sure about that point 4 ? -->

### Images overview

The `debian:testing`-based stack currently includes seven images actively
maintained by the Rocker development team (Table 1).  `r-base` builds on
`debian:testing`, and the other six in the stack each build directly from
`r-base`.  The `r-base` image is unique in that it is designated as the
official image for the R language by the Docker organization itself.
This image is reviewed and built by employees of Docker Inc from the
Dockerfile maintained by the Rocker team.  Consequently, users should
refer to this image in Docker commands without an organization napespace,
e.g. `docker pull r-base` to access the official image.  All other images
in the Rocker project are not individually reviewed and built by Docker
Inc and must be refered to require the organization namespace on Docker
Hub, e.g. `docker pull rocker/r-devel`.

Several of the images in this stack are oriented towards the R development
community. `r-devel`, `drd`, `r-devel-san` & `r-devel-ubsan-clang` all
add a copy of the development version of R side-by-side the current
release of R provided by `r-base`.  On these images, the development
version is aliased to `RD` to distinguish from the current release, `R`.
As the names suggest, each provide slightly different configurations.
Of particular interest are the images providing development R built with
support for C/C++ address sanitizers, which are notoriously difficult
to configure [@edd_sanitizers].

As these images focus on developers and/or as base images for custom uses,
this stack does include many specific R packages. Additional dependencies
and packages can easily be installed from `apt`.  R packages not available
in the `apt` repositories can be installed directly from CRAN using the
`littler` scripts, see *Extending Rocker images* section below.

This stack also include the images `shiny` and `rstudio:testing` that
provide Shiny server and RStudio server IDE from RStudio Inc, built on the
`r-base` image.  RStudio and Shiny are registered trademarks of RStudio
Inc, and their use and the distibution of their software in binary form on
Docker Hub has been granted to the Rocker project by explicit permission
from RStudio. Users should review RStudio's trademark use policy
(<http://www.rstudio.com/about/trademark/>) and address inquiries about
further distribution or other questions to <permissions@rstudio.com>.
The Rocker project also provides images with RStudio server and Shiny
server in the stable versioned stack.

**Build schedule**: The official `r-base` image is rebuilt by Docker
following any updates to the official `debian` images (roughly every
few weeks).  The rest of the stack uses build triggers that rebuild the
images either whenever `r-base` is updated or the Dockerfile sources are
updated on the corresponding GitHub repository.  The only exception in
this stack is the `drd` image, which is rebuilt each week by cron trigger.

image            | description                               | size   | downloads 
---------------- | ----------------------------------------- | ------ | ------- 
[r-base](https://hub.docker.com/r/_/r-base)        | official image with current version of R  | 254 MB | 510,000   
[r-devel](https://hub.docker.com/r/rocker/r-devel) |  R-devel added side-by-side r-base (using alias `RD`)    | 1 GB | 4,000
[drd](https://hub.docker.com/r/rocker/drd) |  lightweight r-devel, built almost daily  | 571 MB | 4,000
[r-devel-san](https://hub.docker.com/r/rocker/r-devel-san)  |  as r-devel, but built with compiler sanatizers | 1.1 GB  | 1,000  
[r-devel-ubsan-clang](https://hub.docker.com/r/rocker/r-devel-ubsan-clang)            |  Sanitizers, clang c compiler (instead of gcc) | 1.1 GB | 525
[rstudio:testing](https://hub.docker.com/r/rocker/r-devel-san)        |  rstudio on debian:testing | 1.1 GB | 1,000 
[shiny](https://hub.docker.com/r/rocker/shiny)     |  shiny-server on r-base               | 409 MB | 63,000

Table: the `debian:testing` image stack

<!-- Should we mention r-apt separately? -->

<!-- Do we describe version tags of r-base, which do provide older versions of R but are based on Dockerfiles which will no longer build since the requested R version is no longer in `unstable`?  -->

### The `debian:stable`-based stack

This stack emphasizes stability and reproducibility of the Docker build.
This stack was introduced much more recently, starting in November
of 2016, in response to considerable user input and requests. The key
feature of this stack is the ability to run older versions of R, along
with the contemporaneous versions of R packages.  A user specifies the
version desired using an image tag, e.g.  `rocker/r-ver:3.3.1` will
refer to an image with R version 3.3.1 installed.  Omitting the tag
is equivalent to using the tag `latest`, which, as the name implies,
will always point to an image using the current R release.  Thus, users
wanting to create downstream Dockerfiles which are based on the current
release at time but will continue to reconstruct the same environment in
the future after newer R versions are released should explicilty include
the corresponding version tag, e.g. `rocker/r-ver:3.4.0` at the time
of writing, and not the `latest` tag.  Users can also run the current
development version of R using the tag `devel`, which is built nightly
from R-devel sources from `subversion`.

**MRAN archives**: To facilitate installation of only contemporaneous
versions of R packages on these images, the default CRAN mirror from which
to install R packages is fixed to a snapshot of CRAN corresponding to
the last date for which that version of R was the latest release. These
snapshots are provided by the MRAN archive created by Revolution
Analytics (now aquired by Microsoft), which archives daily snapshots
of all of CRAN, from which a user can install packages with the usual
`install.packages()` function  [@MRAN]. Users can always override this
default by passing any current CRAN repository explicitly.  Unlike CRAN,
Bioconductor only updates its repositories through bi-annual releases
aligned to R's Spring release schedule. Thus, Bioconductor packages
can be installed in the usual way using `bioclite`, which automatically
selects the Bioconductor release corresponding to the version of R in use.

**Version tags**:  These version tags are propegated throughout this
stack: for instance, `rocker/tidyverse:devel` will provide the currently
released versions of the R packages in the `tidyverse` [@tidyverse]
installed on the nightly build of R-devel.  Developers building packages
on this stack are encouraged to tag their images accordingly as well.
Table 3 indicates which versions of R are currently available in the
stack, going back to `3.1.0`.  While older versions may be added to
the stack at a later date, we note that the MRAN snapshots began in
starting 2014-09-17 and thus go back only to R `3.1` era.  Each tag must
be built from a separate Dockerfile, enabling minor differences in the
build instructions to accomidate changing dependencies.  Dockerfiles for
past versions (e.g. < `3.4.0` currently) are intended to remain static
over the long term, while the tag for the current version, `latest`,
and `devel` may be tweaked to accomidate new features or dependencies.

**Installation**: In this stack, the desired version of R is always built
directly from source rather than the `apt` repositories.  Compilers and
dependencies are still installed from the stable `apt` repositories,
and thus lag behind the more recent versions found in the `testing`
stack. Version tags `3.3.3` and older are based on the Debian 8.0 release,
codenamed `jessie`, while `3.4.0`, `devel`, and `latest` are based on
Debian 9.0, `stretch`, released 2017-06-17, and thus have access to much
newer versions of common system dependencies and compilers.  Dependencies
needed to compile R that are not required at runtime are removed once R is
installed, keeping the base images light-weight for faster download times.
While most system dependencies required by common R packages can still
be installed from the `apt` repositories, occassionally a more recent
version must be compiled from source (e.g. the Gibbs Sampling library,
JAGS, and the geospatial toolkit GDAL, must both be compiled from source
on `debian:jessie` images).  In this stack, users should never install R
packages, using `apt`, since this will install a second, older version
of R from the Debian repositories and a dated version of the R package
(since any `r-cran-pkgname` package in the Debian repositories will
depend on `r-base` in `apt` as well).

**Build schedule**: All images are built automatically from
their corresponding Dockerfiles (found in the GitHub repositories
`rocker-org/rocker-versioned` and `rocker-org/geospatial`). A `cron` job
sends nightly build triggers to Docker Hub to rebuild the `latest` and
`devel` tagged images throughout the stack.  To decrease load on the hub,
build triggers for the numeric version tags are sent monthly. Although
the Dockerfiles for older R versions installs an almost-identical
software environment every time, the monthly rebuilding of these images
on Docker Hub ensures they continue to recieve Debian security updates
from upstream, and proves the build recipe still executes successfully.
Note that rebuilding images with software from external repositories
never produces a bitwise identical image, and thus the image identifier
hash will change at each build.

### Images overview

In this stack, each image builds on the previous image, rather than all
other images building directly on the base image, as in the `testing`
stack. Table 2 lists the names and descriptions of the five images in
this stack, along with image size and approximate download counts from
Docker Hub.  Sizes reflect (compressed) cumulative size: a user who has
already downloaded the most recent version of `r-ver` and then pulls a
copy of `rstudio` image will only need to download the additional 115
MB in the `rstudio` layers and not the full 334 MB listed.  This linear
design limits flexibility (no option for `tidyverse` without `rstudio`)
but simplifies use and maintance.  While no single environment will be
optimal for everyone, both the packages selected in this stack and the
stack ordering reflect considerable community input and tuning.

The `rstudio` image includes a lightweight, easy-to-use and
docker-friendly `init` system, s6 [@s6] for running persistant services,
including the RStudio server.  This system provides a convenient way for
downstream Dockerfile developers to add additional persistent services
(such as an ssh server) to a single container, or additional start-up
or shutdown scripts that should be run when an instance starts up or
shuts down.  The `rstudio` image uses such a start-up script to configure
user settings such as login password and permissions through environmental
variables at run time.  <!-- link wiki howto page? -->

The `tidyverse` image contains all required and suggested dependencies
of the commonly used `tidyverse` and `devtools` R packages, including
external database libraries (e.g. Maria-DB and Postgresql).  Users should
consult the package Dockerfiles or `installed.packages()` list directly
for a complete list of installed packages.  The `verse` library adds
commonly used dependencies, notably a large but not comprehensive LaTeX
environment and java development libraries.  Previously the Rocker
project provided the image `hadleyverse` which was since divided into
`tidyverse` and `verse` through community input.

image            | description                               | size   | downloads 
---------------- | ----------------------------------------- | ------ | ------- 
[r-ver](https://hub.docker.com/r/rocker/r-ver)          |  Version-stable base R & src build tools  | 219 MB | 2,000
[rstudio](https://hub.docker.com/r/rocker/rstudio)      |  Adds rstudio                             | 334 MB | 285,000
[tidyverse](https://hub.docker.com/r/rocker/tidyverse)  |  Adds tidyverse & devtools                | 656 MB | 14,000 (+46,000 as `hadleyverse`)
[verse](https://hub.docker.com/r/rocker/verse)          |  Adds java, tex & publishing-related packages   | 947 MB | 4,000
[geospatial](https://hub.docker.com/r/rocker/geospatial)|  Adds geospatial libraries   |  1.3 GB | 1,000

Table: The `rocker-versioned` stack of images

tag     | apt repos | MRAN date    | Build frequency  | images with tag 
--------|-----------|--------------|------------------|-------------------------------           
devel   | `stretch` | current date | nigthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
latest  | `stretch` | current date | nightly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.4.0   | `stretch` | current date | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.3   | `jessie`  | 2017-04-21   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial` 
3.3.2   | `jessie`  | 2017-03-06   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.1   | `jessie`  | 2016-10-31   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.0   | `jessie`  | 2016-06-21   | monthly          | `r-ver`
3.2.0   | `jessie`  | 2015-06-18   | monthly          | `r-ver`
3.1.0   | `jessie`  | 2014-09-17   | monthly          | `r-ver`

Table: Available tags in the `rocker-versioned` stack.

Several images in the `rocker-versioned` stack can be customized on build
when built locally (rather than pulling prebuilt images from Docker Hub)
by using the `--build-arg` option of `docker build`.   In the `r-ver`
image, users can set `R_VERSION`, `BUILD_DATE` (MRAN default snapshot).
In the `rstudio` image users can set `RSTUDIO_VERSION` (otherwise defaults
to the most recent version), and the `PANDOC_TEMPLATES_VERSION` .

This stack also makes use of Docker metadata labels defined by
<http://schema-label.org>,
indicating image `license` (GPL-2.0), `vcs-url` (GitHub repository), and `vendor` (Rocker Project).  These metadata can be altered or extended in downstream images.

<!--

## Maintenance and Updates

These images are actively maintained.  This means that while an effort is made to preserve the general function of these images over time, both these Dockerfiles and the resulting images are subject to some change over time.  In particular:

- Images are regularly re-built on Docker Hub whenever their base image changes, starting with changes to `debian` Docker image.  This is the rough equivalent of running `apt-get upgrade` on `debian`, since all `apt-get` commands are re-run and will pull in the most current sources.  This allows the images to receive security updates to any packages installed from the `debian` repositories, but will not in general change the versions of any software and is very unlikely to break anything.

- The Dockerfiles themselves are subject to change, to improve performance, ease of use, readability, or other concerns raised in the issues.  These changes should also not alter the general behavior of R or R packages on the image.  These changes can be seen in the git history.  The [rocker-versioned](https://github.com/rocker-org/rocker-versioned) repo will use its own semantic version tagging to indicate changes to this repository, with snapshots from these tags archived on Zenodo.

## License ##

The Dockerfiles in this repository are licensed under the GPL 2 or later.

-->

<!-- 
Use cases: instruction, development, 
-->

## Design principles & use cases

- **Portable**: from laptop to cloud
- **Interfaces**: Console-based (R, bash, etc) or RStudio\textsuperscript{\textregistered} Server, Shiny\textsuperscript{\textregistered} server
- **Sandboxed**: experiment without touching local system
- **Optimized**: community-tweaked configurations
- **Extensible**: interactively and persistent community extensions
- **Versioned**: reproducible environments

### Portability: From laptop to cloud

One common use case for Rocker containers is to provide a fast and
reliable mechanism to deploy a custom R envionment to a remote server, such
as Amazon Web Services Elastic Compute (AWS EC2), DigitalOcean, NSF's
Jetstream servers [@jetstream], or private or institutional server hardware.  Being
able to test code in a predictable, pre-configured R environment on a
local machine and then run the same code in an identical environment on a
remote server (e.g. for access to greater RAM, more processors, or merely
to free up the local machine from a long-running computation) is essential
for low-friction scaling of analysis.  Without such contanerization,
getting code to run appropriately in a remote environment can be a major
undertaking, requiring both time and knowledge many would-be users may
not have.

<!-- comment on use in instructional context -->
This portablility is similarly valuable in an instructional context. Requiring students
to install all necessary software on personal laptops can be particularly challenging
for short workshops, where download and installation time and troubleshooting across
heterogeneous machines can prove time consuming and frustrating for students and instructors alike.
By deploying a Rocker image or Rocker-derived image (see extensibility) on a cloud machine,
an instructor can easily provide all students access to the pre-configured software environment
using only the browser on their laptops.  Unlike a vendor-specific virtual machine image (e.g. Amazon Machine Images, AMIs),
this approach does not preclude students being able to run the same software locally later on.
After installing Docker for their machine they too can run the course-specific software environment
easily on their own laptops and deploy it on other cloud providers.  This strategy has proven
effective in our own experience in both workshops and semester-length courses. Similar Docker-based
cloud deployments have been scaled to courses of 100s of students, e.g. at Duke [@Mine] and
UC Berkeley [@data8].

<!-- NOTE: Not sure if we should include such example commands, may be best left to the documentation? -->

For instance, the following code installs Docker on almost any Linux-based server and
launches a Rocker container providing the RStudio-server environment
over a web interface.

    wget -qO- https://get.docker.com/ | sh 
    sudo docker run -d -p 80:8787 -e PASSWORD=<PICK-A-PASSWORD> rocker/rstudio

Many academic and commerical cloud providers make it
possible to execute such code snippets when an instance is launched, without
ever needing to `ssh` into the machine.
After running these two lines, the user may log into the server merely
by pasting its IP address or DNS name into a browser and entering the
appropriate password.  This provides the user with a familiar,
interactive environment running on a remote machine while requiring
a minimum of expertise.

### Interfaces

An important aspect of the Rocker project design is the ability
for users to interact with the software on the container through either
an interactive shell session (such as the R shell or a bash shell), or
through a web browser accessing the RStudio\textsuperscript{\textregistered} Server
integrated development environment (IDE).  Traditional remote and high-performance
computing workflows for R users have usually required the use of `ssh` and 
a terminal only interface, posing a challenge for interactive graphics 
and a barrier to users unfamiliar with these tools and environments. Accessing
an RStudio instance through the browser removes these barriers.  Rocker
images include the RStudio-server software pre-installed and configured with
the explicit permission of RStudio Inc.

Users of the `rstudio` instances can access a root `bash` shell in an 
instance running the container using

```
docker exec -ti <container-id> bash
```

which can be useful for administrative tasks such as installing system dependencies.
All Rocker images can also be run as an interactive R, RScript or bash shell without running
RStudio, which can be useful for batch jobs or for anyone who prefers that environment.

<!--  Reserve these details for documentation, reference here? --> 
As with any interactive docker instance, users
should specify the interactive (`-i`) and terminal (`-t`, here combined
with interactive as `-ti`) flags, and specify the desired executable
environment (e.g. `R`, though other common options may be `Rscript` or
`bash`):

    docker run --rm -ti rocker/tidyverse R

This example shows the use of the `--rm` flag to indicate that the
container should be removed when the interactive session is finished.

**Linking local volumes**

Users of the Docker's GUI interface, Kitematic who use the
`rstudio`-derived images will automatically have their local kitematic
directory linked to the `/home/rstudio/kitematic` directory on the
Docker container.

<!-- Blah blah blah ... link to documentation instead(?) -->
**Root vs user instances**: By default, all rocker containers run as
root. This is consistent with standard Docker practice, allowing both
interactive users and downstream Dockerfiles to install software directly
without having to switch to root.  The main snag of this approach comes if
a user links a local volume and modifies files on that volume, which will
result in the local file being owned by `root` and not the default user.
To avoid this, specify the non-root user at runtime (username `docker`
on the `r-base` stack, or `rstudio` on images deriving from rstudio)
when running an interactive shell.  For users accessing R through RStudio,
this is not necessary.  The docker container will run as root, but a user
logs in through the RStudio server web interface as the non-root user
"rstudio", and thus any changes made to linked volumes will not alter file
permissions on the home directory.  For more details on user permissions,
see the Rocker wiki.

### Sandboxed

Another feature of Rocker instances is the ability to provide a sandboxed 
environment, isolated from software and potentially from other data on
the machine.  Many users are reluctant to upgrade their suite of installed packages,
which may break their existing code or even their R environment if the
installation goes poorly.  However, upgrading packages and/or the R
environment is often necessary to run analysis of a colleague or access
more recent methods.  Rocker instances offer an easy solution. For 
instance, a user can run R code requiring the most recent versions of R
and related packages inside a Rocker container without having to upgrade their
local installations first.  Conversely, one could use Rocker to run code
on an older R release with prior versions of R packages, again without
having to make any alteration to one's local R install.  Another common
use case is to access an instance of R with support for particular
options such as using gcc or clang compiler sanatizers \citep{edd_sanitizers}.
These requires R itself be built with specialized settings that may not be
not available or familiar to many R users on their native system, but can
be easily deployed by pulling the Rocker images `rocker/r-devel-san` or
`rocker/r-devel-ubsan-clang`.

This sandboxing feature of Docker containers is also valuable in the 
remote computing context, where it can allow system administrators to
provide users with administrative (root) privilages to alter and install
software within a container without altering the computational environment
of other users on the remote machine.  Unlike traditional virtual machines,
these containers do not impose a large footprint of reserved resources --
a typical host can easily support 100s of containers [@docker-vs-vm].

<!-- volume linking isolation -->

### Community Optimized

Having a shared, transparent computational environment created by a publicly hosted, reproducible
recipe facilitates community input into configuration details. R and many of its packages
and related software can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth.  While this flexibility reflects varying needs,
many users rely on default settings which may be optimized more for simplicity of installation
than performance.  The Rocker recipes reflect significant community input on these choices,
as well as the considerable experience and expertise of the R Debian mantainer of 20 years
in what configuration options to use.  This helps create a more finely tuned, optimized reference
implementation of the R environment as well as a platform for comparing and discussing these 
issues which are often overlooked elsewhere.  Issues and Pull Requests on the Rocker 
repositories on GitHub attest to these discussions and improvements.
Widespread use of the Rocker image helps promote both testing of these choices and 
contributions further tweaking the configuration from many members of the R community.

### Versioned

[@Boettiger2015]

### Extensible

Rocker images are designed to provide

Often Rocker images are not deployed directly, but used as a base image
onto which users add software for their own needs in a custom Dockerfile.
We offer a few recommendations here to streamline writing these extensions
and promote best practices.

- littler scripts

 To streamline the installation of R packages in downstream Dockerfiles,
 the `r-base` image provides `littler` package [@littler] and associated
 `install2.r` script, enabling users to install additional packages with
 a command such as:

    install2.r pkg1 pgk2 pkg3 ...

- Dockerfile best practices 
- automated builds, build triggers 
 - version-stable builds

## Conclusions

\bibliography{RJreferences}