### 2017-10-06 manuscript.Rmd
Working through Daniel's suggestions in #2. (still going).
Also, flips the order to present the big picture stuff first and details on the project setup / organization after, as Patrick recommended.

---
title: | 
       | An Introduction to Rocker: 
       | Docker Containers for R
author:
  - name: Carl Boettiger
    affiliation: UC Berkeley 
    address:
    - ESPM Department, University of California,
    - 130 Mulford Hall Berkeley, CA 94720-3114, USA
    email:  cboettig@berkeley.edu
  - name: Dirk Eddelbuettel
    affiliation: Debian and R Projects; Ketchum Trading
    address: 
    - Chicago, IL, USA
    email:  edd@debian.org
abstract: >
  We describe the Rocker project, which provides a widely-used suite of Docker images
  with customized R environments for particular tasks. We discuss how this suite is
  organized, and how these tools can increase portability, scaling, reproducibility,
  and convenience of R users and developers.
preamble: >
   \usepackage{longtable}
output: rticles::rjournal_article
---

<!-- Notes to Dirk: 
     Some of this material may be more appropriate for documentation elsewhere, appearing here in only abbreviated form.  
     To get started, I've just tried to err on the side of comprehensive, particularly since our documentation in places is sparse
     or dated.  Feel free to flag any areas that should be moved in part or whole to documentation.  -->

## Introduction

The Rocker project was launched in October 2014 as a collaboration
between CB & DE to provide high-quality Docker images containing the
R environment \citep{edd2014}. Since that time, the project has seen
both considerable uptake in the community and substantial development
and evolution. Here we seek to document the project's objectives and uses.

### What is Docker?

Docker is a popular open source tool to create, distribute, deploy,
and run software applications using *containers*.  Containers provide a
virtual environment (see \citet{Clark2014} for an overview of common virtual environments) requiring all operating-system components an application needs to run:
code, runtime, system tools, system runtime.  Docker containers are
lightweight as they share the operating system kernel, starting instantly
using a layered filesystem which minimizes disk footprint and download time,
are built on open standards that run on all major platforms (Linux, Mac,
Windows), and provide an added layer of security by running an application
in an isolated environment \citep{what-docker}.  Familiarity with a few key
terms is helpful in understanding this paper. The term "container" refers to 
an isolated software environment on a computer. R users can think of running
a container as analgous to loading an R package; it makes a set of software
functions available.  A Docker "image" is a binary archive of that software,
analgous to an R binary package -- a given version need be downloaded only
once and then can be "run" to create a container whenever it is needed.  
A "Dockerfile" is a recipe, the source-code, to create a Docker image.  Development
and contributions to the Rocker project focuses on the construction, organization
and maintenance of these Dockerfiles.

## Design principles & use cases

<!--  The main point here is that these are generic features that could be goals 
      of any effort to provide researchers with virtual computing environments.

      While most virtual computing environments (AMI images, scientific cloud
      platforms like cyverse, etc) provide some subset of these, they are most
      effective in concert.

- **Portable**: from laptop to cloud
- **Interfaces**: Console-based (R, bash, etc) or RStudio\textsuperscript{\textregistered} Server, Shiny\textsuperscript{\textregistered} server
- **Sandboxed**: experiment without touching local system
- **Transparent**: Dockerfiles & automated builds provide a transparent and verifiable record of precisely what is installed.
- **Optimized**: community-tweaked configurations
- **Extensible**: interactively and persistent community extensions
- **Versioned**: reproducible environments

-->

### Portability: From laptop to cloud

One common use case for Rocker containers is to provide a fast and
reliable mechanism to deploy a custom R environment to a remote server, such
as Amazon Web Services Elastic Compute (AWS EC2), DigitalOcean, NSF's
Jetstream servers \citep{jetstream}, or private or institutional server hardware. 
Lawrence Berkeley National Labs (LBNL) has made `Singularity`: software which can import
and run Docker containers without root privileges, as most cluster administrators require \citep{singularity}.  
Another LBNL project, *Shifter*, allows Docker containers to be deployed on supercomputing
architectures such as CRAY \citep{shifter}.

Importantly, Rocker containers are just as easy to run on locally on most modern laptops
by first installing the appropriate Docker distribution for Windows, MacOS, and Linux-based
operating systems. By sharing volumes with the local host, users can still manipulate
files with familiar, native tools while performing computation through a reproducible,
containerized environment \citep{Boettiger2015}. Being
able to test code in a predictable, pre-configured R environment on a
local machine and to then run the same code in an identical environment on a
remote server (e.g. for access to greater RAM, more processors, or merely
to free up the local machine from a long-running computation) is essential
for low-friction scaling of analysis.  Without such containerization,
getting code to run appropriately in a remote environment can be a major
undertaking, requiring both time and knowledge many would-be users may
not have.

For instance, the following code installs Docker on almost any Linux-based server and
launches a Rocker container providing the RStudio-server environment
over a web interface.

    wget -qO- https://get.docker.com/ | sh 
    sudo docker run -d -p 8787:8787 -e PASSWORD=<PICK-A-PASSWORD> rocker/rstudio

Many academic and commercial cloud providers make it
possible to execute such code snippets when a container is launched, without
ever needing to `ssh` into the machine. The user may log into the server merely
by pasting its IP address or DNS name (followed by the chosen port, RStudio's default `:8787` in the above example)
into a browser and entering the
appropriate password.  This provides the user with a familiar,
interactive environment running on a remote machine while requiring
a minimum of expertise.

This portability is also valuable in an instructional context. Requiring students
to install all necessary software on personal laptops can be particularly challenging
for short workshops, where download and installation time and troubleshooting across
heterogeneous machines can prove time consuming and frustrating for students and instructors alike.
By deploying a Rocker image or Rocker-derived image (see *Extensibility*) on a cloud machine,
an instructor can easily provide all students access to the pre-configured software environment
using only the browser on their laptops. This strategy has proven
effective in our own experience in both workshops and semester-length courses. Similar Docker-based
cloud deployments have been scaled to courses of 100s of students, e.g. at Duke \citep{Mine} and
UC Berkeley \citep{data8}.

### Interfaces

An important aspect of the Rocker project design is the ability
for users to interact with the software on the container through either
an interactive shell session (such as the R shell or a bash shell), or
through a web browser accessing the RStudio\textsuperscript{\textregistered} Server
integrated development environment (IDE).  Traditional remote and high-performance
computing workflows for R users have usually required the use of `ssh` and 
a terminal only interface, posing a challenge for interactive graphics 
and a barrier to users unfamiliar with these tools and environments. Accessing
an RStudio\textsuperscript{\textregistered} container through the browser removes these barriers.  Rocker
images include the RStudio-server software pre-installed and configured with
the explicit permission of RStudio\textsuperscript{\textregistered} Inc.

Users can access a root `bash` shell in a running Rocker container using

```
docker exec -ti <container-id> bash
```

which can be useful for administrative tasks such as installing system dependencies.
All Rocker images can also be run as an interactive R, RScript or bash shell without running
RStudio, which can be useful for batch jobs or for anyone who prefers that environment.

<!--  Reserve these details for documentation, reference here? --> 
As with any interactive Docker container, users
should specify the interactive (`-i`) and terminal (`-t`, here combined
with interactive as `-ti`) flags, and specify the desired executable
environment (e.g. `R`, though other common options may be `Rscript` or
`bash`):

    docker run --rm -ti rocker/tidyverse R

This example shows the use of the `--rm` flag to indicate that the
container should be removed when the interactive session is finished.

Details on sharing volumes, managing user permissions, and more can be found
on the Rocker website, <https://rocker-project.org>.

### Sandboxed

Another feature of Rocker containers is the ability to provide a sandboxed 
environment, isolated from software and potentially from other data on
the machine.  Many users are reluctant to upgrade their suite of installed packages,
which may break their existing code or even their R environment if the
installation goes poorly.  However, upgrading packages and/or the R
environment is often necessary to run analysis of a colleague or access
more recent methods.  Rocker offers an easy solution. For 
instance, a user can run R code requiring the most recent versions of R
and related packages inside a Rocker container without having to upgrade their
local installations first.  Conversely, one could use Rocker to run code
on an older R release with prior versions of R packages, again without
having to make any alteration to one's local R install.  Another common
use case is to access a container with support for particular
options such as using gcc or clang compiler sanitizers \citep{edd_sanitizers}.
These requires R itself be built with specialized settings that may not be
not available or familiar to many R users on their native system, but can
be easily deployed by pulling the Rocker images `rocker/r-devel-san` or
`rocker/r-devel-ubsan-clang`.

This sandboxing feature is also valuable in the 
remote computing context, allowing system administrators to grant
users freedom to install software which requires root privileges inside a container,
while not granting them root access on the host machine. Root access
is required to launch Docker containers, though not to access containers already
running and providing some service such as RStudio. Users logging into a container
through the RStudio\textsuperscript{\textregistered} interface do not by default have root privileges, though are able
to install R packages. Granting these users root privileges in the container still leaves
them sandboxed from the host container.  Further, Docker images can
be deployed from a machine without root privileges through the alternate container
runtime environment, Singularity\citep{singularity}.  Unlike traditional virtual machines,
these containers do not impose a large footprint of reserved resources --
a typical host can easily support 100s of containers \cite{what-docker}.

### Transparent

Users can easily determine the software stack installed on any Rocker image
by examining the associated Dockerfile recipe, which provides a concise, human-readable
record of the installation.
All Rocker images use automated builds through Docker Hub, which also acts as the central,
default repository distributing the images.  Using automated builds rather than uploading
pre-built image binaries to the Docker Hub avoids the potential for the build
not to match the recipe.  The corresponding Dockerfile is visible both on the Docker Hub
and in the linked GitHub repository of the Rocker project (https://github.com/rocker-org), 
which provides a transparent versioned history of all changes made to these recipes,
as well as documentation, a community wiki, and issue trackers for discussing proposed
changes, bugs, improvements to the Dockerfiles and troubleshoot any issues users may encounter.

### Community Optimized

Having a shared, transparent computational environment created by a publicly hosted, reproducible
recipe facilitates community input into configuration details. R and many of its packages
and related software can be configured with a wide range of options, compilers, different
linear-algebra libraries and so forth.  While this flexibility reflects varying needs,
many users rely on default settings which may be optimized more for simplicity of installation
than performance.  The Rocker recipes reflect significant community input on these choices,
as well as the considerable experience and expertise of the R Debian maintainer of 20 years
in what configuration options to use.  This helps create a more finely tuned, optimized reference
implementation of the R environment as well as a platform for comparing and discussing these 
concerns which are often overlooked elsewhere.  Issues and Pull Requests on the Rocker 
repositories on GitHub attest to these discussions and improvements.
Widespread use of the Rocker image helps promote both testing of these choices and 
contributions further tweaking the configuration from many members of the R community.

### Versioned

The versioned stack, (`r-ver`, `rstudio`, `tidyverse`, `verse`,
and `geospatial`), provides images which are intended to build an identical
software stack every time, regardless of the release of new libraries and packages.  This 
can be important for users who need computational reproducibility more than having the latest
release of any piece of software, since subsequent releases can alter the behavior of code,
introducing errors or altering previous results.  Users should specify an R version tag in
the Docker image name to request a version stable image, e.g. `rocker/verse:3.4.0`.  If no tag
is explicitly requested, Docker will provide the image with the tag `:latest`, which will
always have the latest available versions of the software (built nightly).

Users building on the version-tagged images will by default use the MRAN snapshot mirror \citep{MRAN}
associated with the most recent date for which that image was current.  This ensures that
a Dockerfile building `FROM rocker/verse:3.4.0` will only install R package versions that were
available on CRAN on 2017-06-30, that is, the day R v3.4.1 was released.  This default
can of course be overwritten in the standard R manner, e.g. by specifying a different CRAN mirror explicitly in any 
command to `install.packages()`, or by adjusting the default CRAN mirror in `options(repo=<CRAN-MIRROR>)`
in an `.Rprofile`.  Note that the MRAN date associated with the current release (e.g. `3.4.1` at the time of
writing) will continue to advance on the Docker-hub image until the next R release.  Software
installed from `apt-get` in these images will come from the the stable Debian release (`stretch` or `jessie`)
and thus not change versions (though it will receive security patches).  Packages installed from BioConductor
using `bioclite()` utility will also install the version appropriate to the version of R found on the system
(the Bioconductor semi-annual release model avoids the need for an MRAN mirror). Users installing packages
from GitHub or other sources can request a specific git release tag or hash for a more reproducible build,
or adopt an alternative approach such as `packrat` \citep{packrat}. A more general discussion of the use and limitations of Docker for computational reproducibility can be found in \citep{Boettiger2015}. 
<!--
however, `packrat` should not be necessary
for just maintaining reproducibility with respect to CRAN and Bioconductor stacks installed on versioned images.
-->

### Extensible

Any portable computational environment faces an inevitable tension between
the "kitchen sink problem" at one extreme, and the "discovery problem" on
the other.  A kitchen sink image seeks to accommodate too many use cases
in a single image.  Such images are inevitably very large and thus slow
or difficult to deploy, maintain and optimize.  At the other extreme, providing
too many specialized images makes it more difficult for a user to discover 
the one they need.  The Rocker project seeks to avoid both of these problems
by providing a carefully curated suite of images that an be easily extended by
individuals and communities.

To make extensions transparent and persistent, Rocker images can be extended by
any user by writing their own Dockerfiles based on an appropriate Rocker image.  
The Dockerfiles in the Rocker stack should themselves provide a simple of example
of this.  A user begins by selecting an appropriate base image for their needs: 
if the RStudio\textsuperscript{\textregistered} interface is desired, a user might start with `FROM rocker/rstudio`;
an image for testing a particular C code in an R package might use `FROM rocker/r-devel-san`,
an image for reproducing a data analysis will probably select a stable version tag
in addition to an appropriate base library, e.g.: `FROM rocker/tidyverse:3.4.1`
Users can easily add additional software to any running Rocker image using
the standard R and Debian mechanisms. Details on how to extend Rocker images can
be found at <https://rocker-project.org>.

Sharing these Dockerfiles can also facilitate the emergence 
of extensions tuned to particular communities.  For instance, the `rocker/geospatial`
image emerged from the input of a number of Rocker users all adding common
geospatial libraries and packages on top of the existing Rocker images. 
This coalescence helped create a more fine-tuned image with broad support 
for a wide range of commonly used data formats and libraries.  Other community
images are developed and maintained independently of the Rocker project,
such as the `popgen` image of population-genetics-oriented software developed 
by the National Evolutionary Synthesis Center (NESCent). Rocker images are also 
being used as base Docker images in the NSF sponsored Whole Tale project for
reproducible computing \citep{wholetale}, and are heavily used by the R-Hub
project in automated package testing \citep{rhub}.

## Rocker organization and workflow

The Rocker project consists of a suite of images built automatically by
and hosted on the Docker Hub, <https://hub.docker.com/r/rocker>.  Source
Dockerfiles, supporting scripts and documentation are hosted on GitHub
under the organization `rocker-org`, <https://github.com/rocker-org>.
The issue tracker and pull requests are used for community input,
discussions and contributions to these images. The rocker project wiki
provides a place to synthesize community-contributed documentation,
use-cases, and other knowledge about using the Rocker images.

### Images in the Rocker Project

The Rocker project aims to provide a small core of Docker images that
serve as convenient 'base' images on which other users can build custom
R environments by writing their own Dockerfiles, while also providing a
'batteries included' approach of images that can be used out of the box.
The challenges of balancing diverse needs driven by very different use
cases (discussed in more detail below) against the overarching goals
of creating images that are still sufficiently light-weight, easy to use
and easy to maintain is a difficult art.  The implementation in both
individual Rocker images and image stacks can never perfect that balance
for everyone, but today reflects the considerable community input and
testing over past few years.

All Rocker images are based on the Debian Linux distribution.
The Debian platform provides a small base image, the well-known `apt`
package management system and rich ecosystem of software libraries,
making it the base image of choice for Docker images, including many
of the "official" images maintained by Docker's own development team.
The Debian platform is also perhaps the best supported Linux platform
within the R community, including an active `r-sig-debian` listserve.
The relatively long period between stable Debian releases
(roughly two years recently) means that software in the Debian stable 
(e.g. `debian:jessie`, `debian:stretch`)
releases can lag significantly behind current releases of popular
software, including R.  More recent versions of packages can be found in
the pre-release distribution, `debian:testing`, while the very latest
binary builds can be found on `debian:unstable`.   The Rocker project
can be largely divided into two stacks which address different needs,
reflected in which Debian distribution they are based on.  The first stack
is based on `debian:testing`, the second, more recently introduced stack
is based only on Debian stable releases.  Rocker images always point to specific
stable releases (`jessie`, `stretch`), and do not use the tag `debian:stable`,
which is a rolling tag that always points to the most
recent stable version.   The different Rocker stacks have different
aims and thus provide different images, as shown in Tables 1 & 2.

## The `debian:testing`-based images

The `debian:testing` stack aims to makes the most efficient use of
upstream builds -- the precompiled `.deb` binaries provided by the
Debian repositories.  It is both quicker and easier to install software
from binaries, since the package manager (`apt`) manages the necessary
(binary) dependencies and bypasses the time-consuming process of compiling
from source.  Basing this stack on `debian:testing` means that much more
recent versions of commonly-used libraries and compilers are available
as binaries than would be found in a Debian stable release.  In order
to provide access to the most recent available binaries, this stack uses
apt-pinning \citep{apt_pinning} to allow the `apt` package manager to also
install binaries from `debian:unstable`, which represents the most recent,
bleeding edge of packages built for Debian when necessary.  For instance,
the `r-base` image provided by Rocker installs the most recent version of
R as a binary from Debian `unstable`.  Similarly, recent versions of many
popular R packages can also be installed through the package manager,
e.g. `apt-get install r-cran-xml`.  This can be particularly helpful
for packages with external system dependencies (such as `libxml2-dev`
in this example) which cannot be installed from the R console as they 
are system dependencies rather than R packages installed from within R.
We should note, however, that only about 500 of the over
11000 CRAN packages are available as Debian packages.

As the names `testing` and `unstable` imply, this approach is not
without challenges.  The particular version of any given package can
change as packages move from `unstable` into `testing`. New versions
are sent to `unstable` during the normal course of Debian development.
This can occasionally break an previously working installation command
in a Dockerfile until the maintainer redirects the package manager to
install a package from the `unstable` sources that could previously
be installed from `testing`, or vice versa (using the `-t` option in
`apt`)  That said, packages only migrate from 
`unstable` to `testing` after a period of several days---and if the 
migration and installation of the particular version is free of 
interactions with other packages in their dependency graph. That way,
`unstable` serves as validation lab which leaves `testing` reasonably 
stable yet current.

Relative to `stable`, the `testing` stack thus offers the following advantages:

1. These Dockerfiles are easy to develop and extend because almost all software can be installed through the package manager
<!-- DE 1.) would be true of stable too, no?   -->
<!-- CB  those sources are sometimes too old; e.g. jags sources on jessie are incompatible with rjags, etc -->
2. These Dockerfiles always install the most recent available software
3. These Dockerfiles can almost always build relatively quickly

and these down-sides:

1. These Dockerfiles require occasional minor maintenance to ensure successful builds. (e.g. changing an installation directive from `unstable` to `testing` or vice versa)
2. The resulting images are inherently dynamic: rebuilding the same Dockerfile months or years apart will generate images with significantly different versions of software installed.
3. The use of apt pinning may be unfamiliar to some users, where the user must ultimately be responsible to ensure compatibility \citep{apt_pinning}.

### Images overview

The `debian:testing`-based stack currently includes seven images
actively maintained by the Rocker development team (Table 1).
`r-base` builds on `debian:testing`, and the other six in the stack
each build directly from `r-base`.  The `r-base` image is unique in
that it is designated as the official image for the R language by the
Docker organization itself.  This official image is reviewed and then
built by employees of Docker Inc based on a Dockerfile maintained by
the Rocker team.  Consequently, users should refer to this image in
Docker commands without an organization namespace, e.g. `docker run -t
r-base` to access the official image.  All other images in the Rocker
project are not individually reviewed and built by Docker Inc and must
be referenced using the `rocker` namespace,
e.g. `docker run -t rocker/r-devel`.

Several of the images in this stack are oriented towards the R
development community. `r-devel`, `drd`, `r-devel-san` &
`r-devel-ubsan-clang` all add a copy of the development version of R
side-by-side the current release of R provided by `r-base`.  On these
images, the development version is aliased to `RD` to distinguish from
the current release, `R`.  As the names suggest, each provide slightly
different configurations.  Of particular interest are the images
providing development R built with support for C/C++ address and
undefined-behavior sanitizers, which are somewhat more difficult to
configure \citep{edd_sanitizers}.

As these images focus on developers and/or as base images for custom uses,
this stack does not include many specific R packages. Additional dependencies
and packages can easily be installed from `apt`.  R packages not available
in the `apt` repositories can be installed directly from CRAN using the
`littler` scripts, see *Extending Rocker images* section below.

This stack also include the images `shiny` and `rstudio:testing` that
provide Shiny server and RStudio\textsuperscript{\textregistered} server IDE from RStudio\textsuperscript{\textregistered} Inc, built on the
`r-base` image.  RStudio\textsuperscript{\textregistered} and Shiny are registered trademarks of RStudio
Inc, and their use and the distribution of their software in binary form on
Docker Hub has been granted to the Rocker project by explicit permission
from RStudio. Users should review RStudio's trademark use policy
(<http://www.rstudio.com/about/trademark/>) and address inquiries about
further distribution or other questions to <permissions@rstudio.com>.
The Rocker project also provides images with RStudio\textsuperscript{\textregistered} server and Shiny
server in the stable versioned stack.

**Build schedule**: The official `r-base` image is rebuilt by Docker
following any updates to the official `debian` images (roughly every
few weeks).  The rest of the stack uses build triggers that rebuild the
images either whenever `r-base` is updated, or the Dockerfile sources are
updated on the corresponding GitHub repository.  The only exception in
this stack is the `drd` image, which is rebuilt each week by a `cron` trigger.

image            | description                               | size   | downloads 
---------------- | ----------------------------------------- | ------ | ------- 
[r-base](https://hub.docker.com/r/_/r-base)        | official image with current version of R  | 254 MB | 510,000   
[r-devel](https://hub.docker.com/r/rocker/r-devel) |  R-devel added side-by-side r-base (using alias `RD`)    | 1 GB | 4,000
[drd](https://hub.docker.com/r/rocker/drd) |  lightweight r-devel, built almost daily  | 571 MB | 4,000
[r-devel-san](https://hub.docker.com/r/rocker/r-devel-san)  |  as r-devel, but built with compiler sanitizers | 1.1 GB  | 1,000  
[r-devel-ubsan-clang](https://hub.docker.com/r/rocker/r-devel-ubsan-clang)            |  Sanitizers, clang c compiler (instead of gcc) | 1.1 GB | 525
[rstudio:testing](https://hub.docker.com/r/rocker/r-devel-san)        |  rstudio on debian:testing | 1.1 GB | 1,000 
[shiny](https://hub.docker.com/r/rocker/shiny)     |  shiny-server on r-base               | 409 MB | 63,000

Table: The `debian:testing` image stack

### The `debian:stable`-based stack

This stack emphasizes stability and reproducibility of the Docker build.
This stack was introduced much more recently (November 2016) 
in response to considerable user input and requests. The key
feature of this stack is the ability to run older versions of R along
with the then-contemporaneous versions of R packages.  A user specifies the
version desired using an image tag, e.g.  `rocker/r-ver:3.3.1` will
refer to an image with R version 3.3.1 installed.  Omitting the tag
is equivalent to using the tag `latest`, which, as the name implies,
will always point to an image using the current R release.  Thus, users
wanting to create downstream Dockerfiles which are based on the current
release at time (but will continue to reconstruct the same environment in
the future after newer R versions are released) should explicitly include
the corresponding version tag, e.g. `rocker/r-ver:3.4.0` at the time
of writing, and not the `latest` tag.  Users can also run the current
development version of R using the tag `devel`, which is built nightly
from R-devel sources from `subversion`.

**MRAN archives**: To facilitate installation of only contemporaneous
versions of R packages on these images, the default CRAN mirror from which
to install R packages is fixed to a snapshot of CRAN corresponding to
the last date for which that version of R was the latest release. These
snapshots are provided by the MRAN archive created by Revolution
Analytics (now part of Microsoft). It archives daily snapshots
of all of CRAN from which a user can install packages with the usual
`install.packages()` function  \citep{MRAN}. Users can always override this
default by passing any current CRAN repository explicitly.  Unlike CRAN,
Bioconductor only updates its repositories through bi-annual releases
aligned to R's spring release schedule. Thus, Bioconductor packages
can be installed in the usual way using `bioclite`, which automatically
selects the Bioconductor release corresponding to the version of R in use.

**Version tags**:  The version tags are propagated throughout this
stack: _e.g._, `rocker/tidyverse:devel` will provide the currently
released versions of the R packages in the `tidyverse` \citep{tidyverse}
installed on the nightly build of R-devel.  Developers building packages
on this stack are encouraged to tag their images accordingly as well.
Table 3 indicates which versions of R are currently available in the
stack, going back to `3.1.0`.  While older versions may be added to
the stack at a later date, we note that the MRAN snapshots began in
starting 2014-09-17 and thus go back only to the R `3.1` era.  Each tag must
be built from a separate Dockerfile, enabling minor differences in the
build instructions to accommodate changing dependencies.  Dockerfiles for
past versions (e.g. < `3.4.0` currently) are intended to remain static
over the long term, while the tag for the current version, `latest`,
and `devel` may be tweaked to accommodate new features or dependencies.

**Installation**: In this stack, the desired version of R is always built
directly from source rather than the `apt` repositories.  Compilers and
dependencies are still installed from the stable `apt` repositories,
and thus lag behind the more recent versions found in the `testing`
stack. Version tags `3.3.3` and older are based on the Debian 8.0 release,
code-named `jessie`, while `3.4.0`, `devel`, and `latest` are based on
Debian 9.0, `stretch`, released 2017-06-17, and thus have access to much
newer versions of common system dependencies and compilers.  Dependencies
needed to compile R that are not required at runtime are removed once R is
installed, keeping the base images light-weight for faster download times.
While most system dependencies required by common R packages can still
be installed from the `apt` repositories, occasionally a more recent
version must be compiled from source (e.g. the Gibbs Sampling library,
JAGS, and the geospatial toolkit GDAL, must both be compiled from source
on `debian:jessie` images).  In this stack, users should avoid installing R
packages using `apt` without careful consideration, as this will install a second, 
probably different version
of R from the Debian repositories and a dated version of the R package
(since any `r-cran-pkgname` package in the Debian repositories will
depend on `r-base` in `apt` as well).

**Build schedule**: All images are built automatically from
their corresponding Dockerfiles (found in the GitHub repositories
`rocker-org/rocker-versioned` and `rocker-org/geospatial`). A `cron` job
sends nightly build triggers to Docker Hub to rebuild the `latest` and
`devel` tagged images throughout the stack.  To decrease load on the hub,
build triggers for the numeric version tags are sent monthly. Although
the Dockerfiles for older R versions installs an almost-identical
software environment every time, the monthly rebuilding of these images
on Docker Hub ensures they continue to receive Debian security updates
from upstream, and proves the build recipe still executes successfully.
Note that rebuilding images with software from external repositories
never produces a bit-wise identical image, and thus the image identifier
hash will change at each build.

### Images overview

In this stack, each image builds on the previous image, rather than all
other images building directly on the base image, as in the `testing`
stack. Table 2 lists the names and descriptions of the five images in
this stack, along with image size and approximate download counts from
Docker Hub.  Sizes reflect (compressed) cumulative size: a user who has
already downloaded the most recent version of `r-ver` and then pulls a
copy of `rstudio` image will only need to download the additional 115
MB in the `rstudio` layers and not the full 334 MB listed.  This linear
design limits flexibility (no option for `tidyverse` without `rstudio`)
but simplifies use and maintenance.  While no single environment will be
optimal for everyone, both the packages selected in this stack and the
stack ordering reflect considerable community input and tuning.

The `rstudio` image includes a lightweight, easy-to-use and
docker-friendly `init` system, s6 \citep{s6} for running persistent services,
including the RStudio\textsuperscript{\textregistered} server.  This system provides a convenient way for
downstream Dockerfile developers to add additional persistent services
(such as an `ssh` server) to a single container, or additional start-up
or shutdown scripts that should be run when a container starts up or
shuts down.  The `rstudio` image uses such a start-up script to configure
user settings such as login password and permissions through environmental
variables at run time.  <!-- link wiki howto page? -->

The `tidyverse` image contains all required and suggested dependencies
of the commonly-used `tidyverse` and `devtools` R packages, including
external database libraries (e.g. MariaDB and PostgreSQL).  Users should
consult the package Dockerfiles or `installed.packages()` list directly
for a complete list of installed packages.  The `verse` library adds
commonly-used dependencies, notably a large but not comprehensive LaTeX
environment and Java development libraries.  Previously the Rocker
project provided the image `hadleyverse` which was since divided into
`tidyverse` and `verse` through community input.

image            | description                               | size   | downloads 
---------------- | ----------------------------------------- | ------ | ------- 
[r-ver](https://hub.docker.com/r/rocker/r-ver)          |  Version-stable base R & src build tools  | 219 MB | 2,000
[rstudio](https://hub.docker.com/r/rocker/rstudio)      |  Adds rstudio                             | 334 MB | 285,000
[tidyverse](https://hub.docker.com/r/rocker/tidyverse)  |  Adds tidyverse & devtools                | 656 MB | 14,000 (+46,000 as `hadleyverse`)
[verse](https://hub.docker.com/r/rocker/verse)          |  Adds java, tex & publishing-related packages   | 947 MB | 4,000
[geospatial](https://hub.docker.com/r/rocker/geospatial)|  Adds geospatial libraries   |  1.3 GB | 1,000

Table: The `rocker-versioned` stack of images

tag     | apt repos | MRAN date    | Build frequency  | images with tag 
--------|-----------|--------------|------------------|-------------------------------           
devel   | `stretch` | current date | nightly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
latest  | `stretch` | current date | nightly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.4.1   | `stretch` | current date | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.4.0   | `stretch` | current date | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.3   | `jessie`  | 2017-04-21   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial` 
3.3.2   | `jessie`  | 2017-03-06   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.1   | `jessie`  | 2016-10-31   | monthly          | `r-ver`, `rstudio`, `tidyverse`, `verse`, `geospatial`
3.3.0   | `jessie`  | 2016-06-21   | monthly          | `r-ver`
3.2.0   | `jessie`  | 2015-06-18   | monthly          | `r-ver`
3.1.0   | `jessie`  | 2014-09-17   | monthly          | `r-ver`

Table: Available tags in the `rocker-versioned` stack.

Several images in the `rocker-versioned` stack can be customized on build
when built locally (rather than pulling prebuilt images from Docker Hub)
by using the `--build-arg` option of `docker build`.   In the `r-ver`
image, users can set `R_VERSION`, `BUILD_DATE` (MRAN default snapshot).
In the `rstudio` image users can set `RSTUDIO_VERSION` (otherwise defaults
to the most recent version), and the `PANDOC_TEMPLATES_VERSION` .

This stack also makes use of Docker metadata labels defined by
<http://schema-label.org>,
indicating image `license` (GPL-2.0), `vcs-url` (GitHub repository), and `vendor` (Rocker Project).  These metadata can be altered or extended in downstream images.

<!--

## Maintenance and Updates

These images are actively maintained.  This means that while an effort is made to preserve the general function of these images over time, both these Dockerfiles and the resulting images are subject to some change over time.  In particular:

- Images are regularly re-built on Docker Hub whenever their base image changes, starting with changes to `debian` Docker image.  This is the rough equivalent of running `apt-get upgrade` on `debian`, since all `apt-get` commands are re-run and will pull in the most current sources.  This allows the images to receive security updates to any packages installed from the `debian` repositories, but will not in general change the versions of any software and is very unlikely to break anything.

- The Dockerfiles themselves are subject to change, to improve performance, ease of use, readability, or other concerns raised in the issues.  These changes should also not alter the general behavior of R or R packages on the image.  These changes can be seen in the git history.  The [rocker-versioned](https://github.com/rocker-org/rocker-versioned) repo will use its own semantic version tagging to indicate changes to this repository, with snapshots from these tags archived on Zenodo.

## License ##

The Dockerfiles in this repository are licensed under the GPL 2 or later.

-->

<!-- 
Use cases: instruction, development, 
-->

## Conclusions

Over the past several years, Docker has seen immense adoption across industry and academia. 
The Open Container initiative \citep{oci} now provides an open standard that has further
extended this container approach to research environments through projects such as 
Singularity \citep{singularity}, used on clusters without root access, and Shifter \citep{shifter}
for use on leadership high performance computing facilities. Containerization promises
to solve numerous challenges such as portability and replicability in research computing,
which often relies on complex and heterogeneous software stacks \citep{Boettiger2015}. Yet
implementing such environments in containers is not a trivial task, and not all implementations
provide the same usability, portability or reproducibility.  Here we have detailed the 
approach taken by the Rocker project in creating and maintaining these environments through
an open and community-driven process.  This structure of the Rocker project has evolved
over three years of operation while drawing in an ever widening base of academic researchers, 
university instructors and industry users.  We believe this overview will be instructive
not only to users and developers interested in the Rocker project, but as a model for
similar efforts around other environments or domains.

\bibliography{RJreferences}